---
title: 'Machine Learning'
date: "7/5/2022"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction: Please Read

## Data Collection


### Collection:

Handwritten digits are from ten individuals. Each individual is to write each digit (0 - 9) five times (50 written digits). Digits should be written in Black ink (consider a sharpie, black marker, or a highly legible pen) upon white paper (such as computer paper). 

The dataset have 50 copies of every digit (0 - 9), 500 total digits. Digitize each images with a clear camera, and save these into a .ZIP file. 

### 1B) Image Processing

Individual digits are 28x28 grayscale images with pixel values between 0 and 255. A processed MNIST digit is "flattened" into an array of length 784 (28x28). The collection of digits are stored in a Matrix, with each row representing the array of one image. 500 handwritten digits should be processed into one Matrix with dimensions 784x500.

Labels are stored as an array of integers. 500 labels should be stored in one array, length 500.


```{r Q1B}
#Load the image into R
library(imager)
file= "C:/.../1.jpg"
im = load.image(file)
plot(im)


#Convert to grayscale and transpose
im_g = grayscale(im)
plot(im_g)
#Cut the full image into the grids containing individual images
col1 = imsub(im_g,x %inr% c(40,85))
plot(col1)
col2 = imsub(im_g,x %inr% c(150,200))
plot(col2)
col3 = imsub(im_g,x %inr% c(265,310))
plot(col3)
col4 = imsub(im_g,x %inr% c(375,425))
plot(col4)
col5 = imsub(im_g,x %inr% c(490,540))
plot(col5)


#Convert the image into a negative (white space = Black, writing = White)
cut.tuples = list(c(35,90),c(155,210),c(270,325),
                  c(380,440),c(500,560), 
                  c(610,670),c(710,780),c(830,900),
                  c(960,1010),c(1055,1120))

columns = list(col1,col2,col3,col4, col5)

pics = list()

for (i in 1:length(columns)){
  column.vector = columns[[i]]
  for (j in 1:10){
    
    picture = imsub(column.vector,y %inr% c(cut.tuples[[j]][1],cut.tuples[[j]][2]))
    
    pics = append(pics,list(picture))
  }
}

for (i in 1:length(pics)){
  plot(pics[[i]])
}

#Additional processing to make Negative and remove background
pics.background.removed=list()
for (i in 1:length(pics)){
  picture = pics[[i]]<0.6
  pics.background.removed=append(pics.background.removed,list(picture))
}


plot(pics.background.removed[[10]])
#Shrink the image to 28*28
pics.smaller=list()
for (i in 1:length(pics.background.removed)){
  pic.small = resize(pics.background.removed[[i]], size_x = 28, size_y = 28)
  pics.smaller=append(pics.smaller,list(pic.small))
}


for (i in 1:length(pics.smaller)){
  plot(pics.smaller[[i]])
}

#Flatten and load image (as matrices turned vectors), into a matrix with 28*28 columns and 24 rows
flattened = matrix( ncol=28*28)

for (i in 1:length(pics.smaller)){
  flattened = rbind(flattened,c(as.matrix(pics.smaller[[i]])))
}

flattened<-flattened[2:51,]


flattened1=flattened*255
dim(flattened1)

#combine with the data from other images
m = rbind(flattened1, flattened2)


```
